Ore Polynomials in Sage

### 2013-06-10 paper/main.tex

@@

In some cases (for example when the base ring is $\mathbb{Z}$ or
$\mathbb{Z}[x]$), isolated values can be
computed asymptotically faster for large $n$ using the binary splitting
technique.
The \verb|forward_matrix_bsplit| method, called with argument~$n$,
returns a matrix~$P$ and a polynomial~$Q$ such that $P / Q$ multiplied by a
column vector of initial values $c_0, c_1, \ldots$
yields $c_n, c_{n+1}, \ldots$.
This way, computing $10^5$ digits of $\mathrm{e}$ takes a fraction of a second:

\begin{sageexample}
  sage: e_approx = N(e, 400000)
  sage: P, Q = partial_sums.forward_matrix_bsplit(25600)
  sage: u = Matrix([[e_approx], [e_approx]]) - P * Matrix([[1], [2]]) / Q
  sage: u.change_ring(RealField(20))
\end{sageexample}

@@

\subsection{Guessing}

Guessing is, in some sense, the reverse operation of \verb|to_list|. It is one
of the most popular features of packages like gfun, and there are even some
special purpose packages dedicated to this technique~\cite{kauers09a,hebisch11}.
The basic idea is simple. If we are given a finite array of numbers, thought of
as the first terms of an infinite sequence, we may want to know whether this
sequence satisfies a recurrence. The algorithm behind a guessing engine searches
for small equations matching the given data.  Generically, no such equations
exist, so if some are found, it is fair to ``guess'' that they are in fact valid
equations for the whole infinite sequence.

>>>>

\begin{center}
  \begin{tabular}{|l|p{.55\hsize}|}
    \hline
    Method name & short description \\\hline
    \verb|to_D()| & converts annihilator for the coefficients in a power series to a
        differential operator for the sum \\
    \verb|to_F()| & converts shift operator to a difference operator\\
    \verb|to_T()| & converts to a differential operator in terms of the Euler derivative \\
    \verb|annihilator_of_sum()| & converts an annihilator for $f(n)$ to one for $\sum_{k=0}^n f(k)$\\
    \verb|annihilator_of_composition()| & converts an annihilator for $f(n)$ to one for $f(\lfloor un+v \rfloor)$ where $u, v \in \mathbb{Q}$ \\
    \verb|annihilator_of_interlacing()| & interlaces two or more sequences\\
    \verb|generalized_series_solutions()| & computes asymptotic expansions of sequences annihilated by the operator\\ \hline
  \end{tabular}
\end{center}

@@

We compute the asymptotic expansion of the sequence of terms
to estimate how many terms we need to approximate $e$ to a given
number of digits:

\begin{sageexample}
  sage: digits = 10^5
  sage: asymp = inverse_factorials.generalized_series_solutions(3)
  sage: target = lambda x: log(abs(asymp[0](RR(x))), 10) + digits
  sage: num_terms = ceil(find_root(target, 1, 10^6))
  sage: num_terms
\end{sageexample}

In some cases (for example when the base ring is $\mathbb{Z}$ or
$\mathbb{Z}[x]$), isolated values of a sequence can be
computed asymptotically faster for large $n$ than by listing all
values, using the binary splitting technique.
The \verb|forward_matrix_bsplit| method, called with argument~$n$,
returns a matrix $P$ and a denominator $Q$ such that $P / Q$ multiplied a
column vector of initial values $c_0, c_1, \ldots$
yields $c_n, c_{n+1}, \ldots$.
This way, computing $10^5$ digits of $\mathrm{e}$ takes a fraction of a second:

\begin{sageexample}
  sage: e_approx = N(e, 400000)
  sage: P, Q = partial_sums.forward_matrix_bsplit(num_terms)
  sage: u = Matrix([[e_approx], [e_approx]]) - P * Matrix([[1], [2]]) / Q
  sage: u.change_ring(RealField(20))
\end{sageexample}

@@

\subsection{The $q$-Case}

J and Q

\section{Guessing}\label{sec:5}

[[CITE RUBEY.]]

Guessing is one of the most popular features of software packages like gfun.  It
is a technique for detecting patterns in sequences of numbers, which is often
used in combinatorics. The basic idea is simple. If we are given a finite array
of numbers, though of as the first terms of an infinite sequence, we may want to
know whether this sequence satisfies a recurrence. The algorithm behind a
guessing engine searches for small equations matching the given data.
Generically, no such equations exist, so if some are found, it is fair to
``guess'' that they are in fact valid equations for the whole infinite sequence.
=======
\subsection{Guessing}

Guessing is, in some sense, the reverse operation of \verb|to_list|. It is one
of the most popular features of packages like gfun, and there are even some
special purpose packages dedicated to this technique~\cite{kauers09a,hebisch11}.
The basic idea is simple. If we are given a finite array of numbers, thought of
as the first terms of an infinite sequence, we may want to know whether this
sequence satisfies a recurrence. The algorithm behind a guessing engine searches
for small equations matching the given data.  Generically, no such equations
exist, so if some are found, it is fair to ``guess'' that they are in fact valid
equations for the whole infinite sequence.
>>>>>>> manuel/master
